A Seamless, Client-Centric Programming Model for Type Safe Web Applications

### 2014-06-25 haskell10f-ekblad.tex
Make code horribly ugly to fix overflow in PS.

\label{sec:jsexample}
\begin{listingfloat}
\begin{lstlisting}[language=JavaScript]
function handshake(sock) {sock.send('helo');}
function chat(sock, msg) {sock.send('text' + msg);}

window.onload = function() {
  var logbox = document.getElementById('log');
  var msgbox = document.getElementById('message');
  var sock = new WebSocket('ws://example.com');

  sock.onmessage = function(e) {
    logbox.value = e.data + NEWLINE + logbox.value;
  };

  sock.onopen = function(e) {
    handshake(sock);
    msgbox.addEventListener('keydown', function(e) {
      if(e.keyCode == 13) {
        var msg = msgbox.value;
        msgbox.value = '';
        chat(sock, msg);
      }
    });
  };
};
\end{lstlisting}
\caption{JavaScript chatbox implementation}
\label{lst:javascript-client}
\end{listingfloat}

Since the ``chatbox'' application is very simple - users should only be able to
send and receive text messages in real time - we opt for a very simple
design. Two UI elements, \lstinline!logbox! and \lstinline!msgbox!, represent
the chat log and the text area where the user inputs their messages
respectively. When a message arrives, it is prepended to the chat log, making
the most recent message appear at the top of the log window, and when the user
hits the return key in the input text box the message contained therein is sent
and the input text box is cleared.

@@

  runClient $ do
    name <- prompt "Hi there, what is your name?"
    onServer (greetings <.> name)
\end{code}

@@

\begin{code}
setup :: IO ()
setup =
  runApp (mkConfig "ws://localhost:1111" 1111) main
\end{code}

@@

  count <- remote $ do
    r <- remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1, v+1))

@@

  runClient $ do
    withElems ["log","message"] $ \[log,msgbox] -> do
      onServer hello
\end{code}

@@

The \lstinline!withElems! function is part of the Haste compiler's bundled DOM
manipulation library; it locates references to the DOM nodes with the given
identifiers and passes said references to a function.
In this case the variable \lstinline!log! will be bound to the node
with the identifier ``log'', and \lstinline!msgbox! will be bound to the node
identified by ``message''. These are the same DOM nodes that were referenced in
our original example, and refer to the chat log window and the text input
field respectively. After locating all the needed UI elements, the client
proceeds to register itself with the server's recipient list using the
\lstinline!hello! remote computation.

\begin{code}
      let recvLoop chatlines = do
            setProp log "value" $ unlines chatlines
            message <- onServer awaitMsg
            recvLoop (message : chatlines)
      fork $ recvLoop []
\end{code}

@@

\begin{code}
      msgbox `onEvent` OnKeyPress $ \13 -> do
        msg <- getProp msgbox "value"
        setProp msgbox "value" ""
        onServer (sendMsg <.> msg)
\end{code}

@@

\begin{code}
newtype Server a = Server (ReaderT SessionInfo IO a)
  deriving (Functor, Applicative, Monad, MonadIO)
data Remote a = RemoteDummy
\end{code}

@@

\begin{code}
remote f = App $ do
  (next_id, remotes) <- get
  put (next_id+1, (next_id, mkRemote f) : remotes)
  return RemoteDummy
\end{code}

@@

\begin{code}
onEvent :: [(CallID, Method)] -> JSON -> IO ()
onEvent mapping incoming = do
  let (nonce, identifier, args) = fromJSON incoming
      Just f = lookup identifier mapping
  result <- f args
  webSocketSend $ toJSON (nonce, result)
\end{code}

@@

\begin{code}
type Nonce = Int
type ClientState = (Nonce, Map Nonce (MVar JSON))
type Client = StateT ClientState Conc
\end{code}

@@

-- A MutableVar is completely opaque to Haskell code
-- and is only ever manipulated in JavaScript. Thus,
-- we use the Unpacked type to represent it,
-- indicating a completely opaque value.
newtype MutableVar a = MV Unpacked

@@

newMutable :: Marshal a => a -> IO (MutableVar a)
newMutable = ffi "(function(x) {return {val: x};})"

setMutable :: Marshal a => MutableVar a -> a -> IO ()
setMutable = ffi "(function(m, x) {m.val = x;})"

getMutable :: Marshal a => MutableVar a -> IO a
getMutable = ffi "(function(m) {return m.val;})"
\end{code}
\caption{Mutable variables with \lstinline!Haste.Foreign!}
\label{lst:ffi}
\end{listingfloat}

>>>>

\label{sec:jsexample}
\begin{listingfloat}
\begin{lstlisting}[language=JavaScript]
function handshake(sock) {sock.send('helo');}
function chat(sock,msg) {sock.send('text'+msg);}

window.onload = function() {
  var log = document.getElementById('log');
  var msg = document.getElementById('msg');
  var sock = new WebSocket('ws://example.com');

  sock.onmessage = function(e) {
    log.value = e.data + NEWLINE + log.value;
  };

  sock.onopen = function(e) {
    handshake(sock);
    msg.addEventListener('keydown', function(e) {
      if(e.keyCode == 13) {
        var msg = msg.value;
        msg.value = '';
        chat(sock, msg);
      }
    });
  };
};
\end{lstlisting}
\caption{JavaScript chatbox implementation}
\label{lst:javascript-client}
\end{listingfloat}

Since the ``chatbox'' application is very simple - users should only be able to
send and receive text messages in real time - we opt for a very simple
design. Two UI elements, \lstinline!log! and \lstinline!msg!, represent
the chat log and the text area where the user inputs their messages
respectively. When a message arrives, it is prepended to the chat log, making
the most recent message appear at the top of the log window, and when the user
hits the return key in the input text box the message contained therein is sent
and the input text box is cleared.

@@

  runClient $ do
    name <- prompt "Hello, what is your name?"
    onServer (greetings <.> name)
\end{code}

@@

\begin{code}
setup :: IO ()
setup = runApp cfg main
  where cfg = mkConfig "ws://localhost:1111" 1111
\end{code}

@@

  count <- remote $ do
    r <- remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1,v+1))

@@

  runClient $ do
    withElems ["log","msg"] $ \[log,msg] -> do
      onServer hello
\end{code}

@@

The \lstinline!withElems! function is part of the Haste compiler's bundled DOM
manipulation library; it locates references to the DOM nodes with the given
identifiers and passes said references to a function.
In this case the variable \lstinline!log! will be bound to the node
with the identifier ``log'', and \lstinline!msg! will be bound to the node
identified by ``msg''. These are the same DOM nodes that were referenced in
our original example, and refer to the chat log window and the text input
field respectively. After locating all the needed UI elements, the client
proceeds to register itself with the server's recipient list using the
\lstinline!hello! remote computation.

\begin{code}
      let recvLoop hist = do
            setProp log "value" $ unlines hist
            message <- onServer awaitMsg
            recvLoop (message : hist)
      fork $ recvLoop []
\end{code}

@@

\begin{code}
      msg `onEvent` OnKeyPress $ \13 -> do
        msgtext <- getProp msg "value"
        setProp msg "value" ""
        onServer (sendMsg <.> msgtext)
\end{code}

@@

\begin{code}
newtype Server a =
  Server (ReaderT SessionInfo IO a)
  deriving (Functor,Applicative,Monad,MonadIO)
data Remote a = RemoteDummy
\end{code}

@@

\begin{code}
remote f = App $ do
  (next_id, remotes) <- get
  put (next_id+1, (next_id,mkRemote f):remotes)
  return RemoteDummy
\end{code}

@@

\begin{code}
onEvent :: [(CallID, Method)] -> JSON -> IO ()
onEvent mapping json = do
  let (nonce, ident, args) = fromJSON json
      Just f = lookup ident mapping
  result <- f args
  webSocketSend $ toJSON (nonce, result)
\end{code}

@@

\begin{code}
type Nonce = Int
type ClientState = (Nonce,Map Nonce (MVar JSON))
type Client = StateT ClientState Conc
\end{code}

@@

-- A MutableVar is completely opaque to Haskell
-- code and is only ever manipulated in
-- JavaScript. Thus, we use the Unpacked type
-- to represent it, indicating a completely
-- opaque value.
newtype MutableVar a = MV Unpacked

@@

newMutable :: Marshal a =>
              a -> IO (MutableVar a)
newMutable = ffi "(function(x){return {val:x};})"

setMutable :: Marshal a =>
              MutableVar a -> a -> IO ()
setMutable = ffi "(function(m, x){m.val = x;})"

getMutable :: Marshal a =>
              MutableVar a -> IO a
getMutable = ffi "(function(m){return m.val;})"
\end{code}
\caption{Mutable variables with \lstinline!Haste.Foreign!}
\label{lst:ffi}
\end{listingfloat}
